Management induced changes of soil organic carbon on global croplands

### 2020-12-17 SOCbudgetPaper/conclusion.Rmd
update (fix citations, affilations, figure position, LPJmL4Paper to LPJmL4, sidewaytable)

Our global SOC model is able to estimate spatially explicit SOC stocks, SOC debts and stock change factor considering agricultural management. It is -- to our knowledge -- the first study that quantifies the impact of time-variant and spatially explicit historic agricultural management on global SOC stocks.

Our results clearly demonstrate that agricultural management needs to be explicitly considered in global carbon assessments and models. That also implies that we need better monitoring of agricultural practices to create this data, but also better accessibility of existing data. Our data-set and the MADRaT package ([@dietrich_madrat_2020]) constitute a starting point for building comprehensive data sets on agricultural management aspects.

Our study again highlights that the expansion of croplands is still a major source of CO~2~ emissions -- not only by the removal of vegetation, but also by a slow depletion of soils. Our estimates indicate a SOC debt of 22 GtC in 2010, and every additional deforestation adds to this debt.

@@

Continuing the historical development, a further closure of the yield gaps may be beneficial to SOC stocks. Forest protection schemes for climate change mitigation would not only reduce SOC losses on newly converted land, they would likely also require further productivity improvements in existing croplands to meet crop demand ([@popp_land-use_2014-1]) -- and thereby usually further reduce the SOC debt on croplands.

@@

Yet, there is ample potential for further improved SOC management. As shown in Fig. \@ref(fig:FlowFig), approximately one fifth of total annual C sequestration by crops is lost through soils (0.8 GtC per year). Large losses in fact occur at the end of the food supply chain (1.2 GtC year), and at the litter soil barrier (1.4 GtC). Improved management could include, firstly, a circular flow from the food supply chain back to soils. Waste composting or excreta recycling could represent a major additional N input to cropland soils ([@brenzinger_organic_2018]). Secondly, soil carbon sequestration techniques ([@smith_soil_2016]), deep ploughing ([@alcantara_deep_2016]) or the transformation of c inputs to more recalcitrant biochar ([@woolf_sustainable_2010]) may transfer larger parts of the biomass at the litter soil barrier into permanent soil pools. Thirdly, reducing the share of residue burning and improved manure recycling could further increase C inputs. Finally, other carbon-accumulating practices, such as the cultivation of cover crops ([@poeplau_carbon_2015]) and agroforestry ([@lorenz_soil_2014]) could increase total C sequestration on croplands.

>>>>

Our global SOC model is able to estimate spatially explicit SOC stocks, SOC debts and stock change factor considering agricultural management. It is --- to our knowledge --- the first study that quantifies the impact of time-variant and spatially explicit historic agricultural management on global SOC stocks.

Our results clearly demonstrate that agricultural management needs to be explicitly considered in global carbon assessments and models. That also implies that we need better monitoring of agricultural practices to create this data, but also better accessibility of existing data. Our data-set and the MADRaT package [@dietrich_madrat_2020] constitute a starting point for building comprehensive data sets on agricultural management aspects.

Our study again highlights that the expansion of croplands is still a major source of CO~2~ emissions --- not only by the removal of vegetation, but also by a slow depletion of soils. Our estimates indicate a SOC debt of 22 GtC in 2010, and every additional deforestation adds to this debt.

@@

Continuing the historical development, a further closure of the yield gaps may be beneficial to SOC stocks. Forest protection schemes for climate change mitigation would not only reduce SOC losses on newly converted land, they would likely also require further productivity improvements in existing croplands to meet crop demand [@popp_land-use_2014-1] --- and thereby usually further reduce the SOC debt on croplands.

@@

Yet, there is ample potential for further improved SOC management. As shown in Fig. \@ref(fig:FlowFig), approximately one fifth of total annual C sequestration by crops is lost through soils (0.8 GtC per year). Large losses in fact occur at the end of the food supply chain (1.2 GtC year), and at the litter soil barrier (1.4 GtC). Improved management could include, firstly, a circular flow from the food supply chain back to soils. Waste composting or excreta recycling could represent a major additional N input to cropland soils [@brenzinger_organic_2018]. Secondly, soil carbon sequestration techniques [@smith_soil_2016], deep ploughing [@alcantara_deep_2016] or the transformation of c inputs to more recalcitrant biochar [@woolf_sustainable_2010] may transfer larger parts of the biomass at the litter soil barrier into permanent soil pools. Thirdly, reducing the share of residue burning and improved manure recycling could further increase C inputs. Finally, other carbon-accumulating practices, such as the cultivation of cover crops [@poeplau_carbon_2015] and agroforestry [@lorenz_soil_2014] could increase total C sequestration on croplands.

### 2020-12-17 SOCbudgetPaper/discussion.Rmd
update (fix citations, affilations, figure position, LPJmL4Paper to LPJmL4, sidewaytable)

The evaluation of SOC stock gains and losses is complex and has several dimensions as climatic and anthropogenic effects overlap. If defining the SOC debt of 1975 as the baseline, and measuring land-use emissions on croplands as the difference between a potential natural state and the state under human interventions (see [@pugh_simulated_2015]), global croplands have acted as a carbon sink since 1975 according to our study. However, annual C sequesteration rates of 0.2 per 1000 are well below the promoted 4 per 1000 ([@minasny_soil_2017]), indicating that productivity gains on historic levels alone are not enough to meet ambitious climate targets.

According to [@sanderman_soil_2017], the SOC debt since the beginning of human cropping activities has been at around 37 GtC for the first 30 cm of the soil with half of it attributed to SOC depletion on grasslands. Our estimate of 22 GtC in 2010 for cropland debt is higher as [@sanderman_soil_2017] estimations. However, there are large uncertainties in modeling SOC at the global scale, and [@sanderman_soil_2017] pointed out that their results might be conservatively low compared to experimental results. This suggests that our results are within a plausible range.

Furthermore, [@sanderman_soil_2017] modeled historic trends based on agricultural land expansion without considering SOC variations due to time-variant agricultural management. [@pugh_simulated_2015] considered management effects like tillage and residue returning in a static way, but neither changes over time nor alignment to observed historic data like yield-levels or no-tillage areas were taken into account. Their study moreover concludes that crop productivity-gains (increasing yield levels by 18% in their simulations) do not lead to a substantial decline in SOC debt (less than 1% change). Historic productivity increases were, however, notably larger. Despite large spatial heterogeneity aggregate yield increase rates are often estimated to be well above 50% ([@pellegrini_crop_2018], [@ray_recent_2012], [@rudel_agricultural_2009]).

Our study for the first time uses a dynamic management dataset as driver for SOC dynamics. We show that the moderate global cropland expansion of around 11% between 1974 and 2010 and the resulting depletion of SOC stocks in converted cropland has been outweighed by improvements in agricultural productivity and practices. This is in contrast to [@pugh_simulated_2015] findings of only small effects due to improved practices.

Looking on historic cropland intensification rates of above 50% ([@rudel_agricultural_2009]), a large fraction of increased C input from residue biomass is attributed to productivity improvements.

@@

Modeling management effects at the global scale comes with parametric and structural uncertainties. High SOC gains in the first 30 cm of the soil compared to natural vegetation in e.g. the temperate croplands of Central Europe and the United Kingdom indicate suspiciously high estimates of SOC inputs. As pointed out by [@keel_large_2017] and [@smith_how_2020], carbon input calculations are highly sensitive to the choice of allometric functions determining below- and above-ground residue estimates from harvested quantities (see \@ref(tab:c2dm) for coefficients used in this study). [@keel_large_2017] question whether below ground residues might increase with a fixed root:shoot ratio rather than being independent of productivity gains. Moreover, the study pointed out that plant breeding shifts allometries, which might not be reflected in outdated data sources. While our study considers a dynamic harvest index with rising yields for several crops, we may still overestimate residue biomass in particular for below-ground biomass. However, looking on the evaluation of management effects (see Sect. \@ref(sec:ipcccompare)), there is no general indication of overestimating stock change factors at least compared to IPCC default assumptions. Moreover, it is also likely that we are still missing carbon inputs to the soil from e.g. cover- and inter-cropping practices.

Another uncertainty is connected with the initialization of SOC stocks in 1901, which is assumed to be in steady state considering the land-use pattern of 1901 and agricultural management data of 1965. As shown in Fig. \@ref(fig:SOCinit) the SOC debt estimate almost halves (from ~26 GtC to ~14 GtC), and the SOC debt reduction is strongly reduced (from ~4 GtC to ~2.5 GtC), if considering initialization SOC stocks under undisturbed natural vegetation. [@pugh_simulated_2015] pointed out the importance of accounting for the land-use history, as many CO~2~ emissions from agricultural soils are caused by historic land-use change (LUC) and the slow decline of SOC under cropland before it reaches a new equilibrium. Our results of the Initial-natveg scenario show that the qualitative finding of a reduction of SOC debt through improved agricultural management is robust to changing the initialisation of soil organic carbon, even though the level of SOC debt is sensitive to the initialization setting.

Generally, the limit to the first 30 cm of the soil profile follows the IPCC guidelines ([@eggleston_ipcc_2006], [@calvo_buendia_ipcc_2019]) and assumes that most of the SOC dynamic happens in the topsoil. In this regard several aspects are strongly simplified within our approach. Firstly, distribution of carbon inputs into different soil layers are neglected and all carbon inputs are allocated to the topsoil. This particularly overestimates SOC stocks in the first 30 cm of soil below deeper rooting vegetation, which is certainly the case for most of the woody natural vegetated areas. Secondly changes to the subsoil due to tillage are neglected. As [@powlson_limited_2014] have shown, the subsoil can be a game changer in evaluating total SOC losses or gains for no-tillage systems. No-tillage effects may seem larger than the actually are, if only focusing on the topsoil. SOC transfer to deeper soil layers under tillage, might enhance subsoil SOC compared to no-till practices. Thirdly, organic soils (like peat- and wetlands) and drained cropland areas are not explicitly considered and emissions from these cropland areas are thus likely substantially underestimated.

This study excludes not only peatland degradation, it also does not account for carbon displacement via leaching and erosion. However, as pointed out by [@doetterl_erosion_2016], the final fate of leached or eroded carbon is uncertain and might even offset LUC emissions ([@wang_human-induced_2017]). Whereas for soil quality analysis SOC displacement might play an important role, in this budget approach focusing especially on the SOC debt, displaced but not emitted SOC can be treated as SOC that remains on the cropland.
Moreover, the exclusive focus on croplands ignores LUC emission on other land-use types such as pastures, rangelands and forestry. Human interventions have led to large changes in SOC stocks there as well ([@sanderman_soil_2017],[@friedlingstein_global_2019]). This study does not intend to be a comprehensive LUC emission analysis and acknowledges that land-use changes comes with large overall emissions.

@@

To validate our modeled SOC stocks and stock changes under management, we compare our results to default IPCC stock changes factors of 2006 ([@eggleston_ipcc_2006] and their refinements in 2019 [@calvo_buendia_ipcc_2019]). Both estimates are based on measurement data for croplands (see Table \@ref(tab:SCFglo)). To allow for comparison, we aggregate our stock change factors to the four IPCC climate zones (Fig. \@ref(fig:CLIMzone)).

```{r SCFglo, echo=FALSE, results = 'asis'}
table <- readRDS("../ResultNotebooks/Output/SFC_comparison.rds")
library(xtable)
print(xtable(table$comparison_table, digit=2, comment=FALSE, label="tab:SCFglo", 
              caption="$F^{\\mathrm{SCF}}$ in comparison to IPCC Tier 1 default factors: Stock change factors are in good agreement with the default values of the IPCC in general. For the tropical regions the assumptions changed notablly from the guidelines in 2006 to the update in 2019. leaving our results in very good agreement with the old default assumptions. Default assumption are given under the assumption of medium input systems, which, considering the yield gap in mainly developing regions in the tropics, might be an overestimation and decrease $F^{\\mathrm{SCF}}$ by additional 5-8 percent. Also modelled $F^{SOC}$ have increased for all climates over time."))
```

Whereas our estimates are lower in the two tropical climate zones, temperate zone default factors are higher than [@calvo_buendia_ipcc_2019]. The tropical factors differ substantially between the IPCC guidelines in 2006 ([@eggleston_ipcc_2006]) and their refinement in 2019 ([@calvo_buendia_ipcc_2019]), with our estimates within this range.
Taking the simplified assumption, that tropical soils might suffer from insufficient C input rates (low) due to yield gaps, whereas temperate soils in developed regions might be highly managed and fertilized (high), the difference vanishes partially. Our results still show  especially in temperate dry regions -- the smallest region area-wise -- small deviations from natural SOC stocks. Considering the impact of irrigation and fertilization on carbon-poor dryland soils, even factors above 1 (see Fig. \@ref(fig:SOCmaps)(c)) may be expected.

@@

The world's SOC stock and its changes are highly uncertain, which is seen in the wide range of global SOC stock estimates for the first 30 cm of the soil profile ([@batjes_harmonized_2016], [@hengl_soilgrids250m_2017], [@fao_global_2018], [@schaphoff_lpjml4_2018-1]) in Table \@ref(fig:SOCglo)).

@@

The global estimates of the total SOC stock from this study are on the lower end compared to other modeled results or data driven estimates. SoilGrids ([@hengl_soilgrids250m_2017]) especially stands out for their high estimation, whereas all other sources (including our study) are comparabely similar. Looking at regional results in Fig. \@ref(fig:SOCreg), our estimates turn out to be in good agreement for most of the world, with the largest deviations for boreal moist and tropical moist areas. To avoid that this bias influences our results, which originates from uncertainties in the representation of natural land, we focus on SOC changes on cropland. Pristine natural vegetated areas (like permafrost and rain forests) without human land management thus drop out in our calculation of SOC debt and stock change factors.

>>>>

The evaluation of SOC stock gains and losses is complex and has several dimensions as climatic and anthropogenic effects overlap. If defining the SOC debt of 1975 as the baseline, and measuring land-use emissions on croplands as the difference between a potential natural state and the state under human interventions [see @pugh_simulated_2015], global croplands have acted as a carbon sink since 1975 according to our study. However, annual C sequesteration rates of 0.2 per 1000 are well below the promoted 4 per 1000 [@minasny_soil_2017], indicating that productivity gains on historic levels alone are not enough to meet ambitious climate targets.

According to Sanderman et al. [-@sanderman_soil_2017], the SOC debt since the beginning of human cropping activities has been at around 37 GtC for the first 30 cm of the soil with half of it attributed to SOC depletion on grasslands. Our estimate of 22 GtC in 2010 for cropland debt is higher as Sanderman et al. [-@sanderman_soil_2017] estimations. However, there are large uncertainties in modeling SOC at the global scale, and Sanderman et al. [-@sanderman_soil_2017] pointed out that their results might be conservatively low compared to experimental results. This suggests that our results are within a plausible range.

Furthermore, Sanderman et al. [-@sanderman_soil_2017] modeled historic trends based on agricultural land expansion without considering SOC variations due to time-variant agricultural management. Pugh et al. [-@pugh_simulated_2015] considered management effects like tillage and residue returning in a static way, but neither changes over time nor alignment to observed historic data like yield-levels or no-tillage areas were taken into account. Their study moreover concludes that crop productivity-gains (increasing yield levels by 18% in their simulations) do not lead to a substantial decline in SOC debt (less than 1% change). Historic productivity increases were, however, notably larger. Despite large spatial heterogeneity aggregate yield increase rates are often estimated to be well above 50% [@pellegrini_crop_2018; @ray_recent_2012; @rudel_agricultural_2009].

Our study for the first time uses a dynamic management dataset as driver for SOC dynamics. We show that the moderate global cropland expansion of around 11% between 1974 and 2010 and the resulting depletion of SOC stocks in converted cropland has been outweighed by improvements in agricultural productivity and practices. This is in contrast to Pugh et al. [-@pugh_simulated_2015] findings of only small effects due to improved practices.

Looking on historic cropland intensification rates of above 50% [@rudel_agricultural_2009], a large fraction of increased C input from residue biomass is attributed to productivity improvements.

@@

Modeling management effects at the global scale comes with parametric and structural uncertainties. High SOC gains in the first 30 cm of the soil compared to natural vegetation in e.g. the temperate croplands of Central Europe and the United Kingdom indicate suspiciously high estimates of SOC inputs. As pointed out by Keel et al. [-@keel_large_2017] and Smith et al. [-@smith_how_2020], carbon input calculations are highly sensitive to the choice of allometric functions determining below- and above-ground residue estimates from harvested quantities (see \@ref(tab:c2dm) for coefficients used in this study). Keel et al. [-@keel_large_2017] question whether below ground residues might increase with a fixed root:shoot ratio rather than being independent of productivity gains. Moreover, the study pointed out that plant breeding shifts allometries, which might not be reflected in outdated data sources. While our study considers a dynamic harvest index with rising yields for several crops, we may still overestimate residue biomass in particular for below-ground biomass. However, looking on the evaluation of management effects (see Sect. \@ref(sec:ipcccompare)), there is no general indication of overestimating stock change factors at least compared to IPCC default assumptions. Moreover, it is also likely that we are still missing carbon inputs to the soil from e.g. cover- and inter-cropping practices.

Another uncertainty is connected with the initialization of SOC stocks in 1901, which is assumed to be in steady state considering the land-use pattern of 1901 and agricultural management data of 1965. As shown in Fig. \@ref(fig:SOCinit) the SOC debt estimate almost halves (from ~26 GtC to ~14 GtC), and the SOC debt reduction is strongly reduced (from ~4 GtC to ~2.5 GtC), if considering initialization SOC stocks under undisturbed natural vegetation. Pugh et al. [-@pugh_simulated_2015] pointed out the importance of accounting for the land-use history, as many CO~2~ emissions from agricultural soils are caused by historic land-use change (LUC) and the slow decline of SOC under cropland before it reaches a new equilibrium. Our results of the Initial-natveg scenario show that the qualitative finding of a reduction of SOC debt through improved agricultural management is robust to changing the initialisation of soil organic carbon, even though the level of SOC debt is sensitive to the initialization setting.

Generally, the limit to the first 30 cm of the soil profile follows the IPCC guidelines [@eggleston_ipcc_2006; @calvo_buendia_ipcc_2019] and assumes that most of the SOC dynamic happens in the topsoil. In this regard several aspects are strongly simplified within our approach. Firstly, distribution of carbon inputs into different soil layers are neglected and all carbon inputs are allocated to the topsoil. This particularly overestimates SOC stocks in the first 30 cm of soil below deeper rooting vegetation, which is certainly the case for most of the woody natural vegetated areas. Secondly changes to the subsoil due to tillage are neglected. As Powlson et al. [-@powlson_limited_2014] have shown, the subsoil can be a game changer in evaluating total SOC losses or gains for no-tillage systems. No-tillage effects may seem larger than the actually are, if only focusing on the topsoil. SOC transfer to deeper soil layers under tillage, might enhance subsoil SOC compared to no-till practices. Thirdly, organic soils (like peat- and wetlands) and drained cropland areas are not explicitly considered and emissions from these cropland areas are thus likely substantially underestimated.

This study excludes not only peatland degradation, it also does not account for carbon displacement via leaching and erosion. However, as pointed out by Doetterl et al. [-@doetterl_erosion_2016], the final fate of leached or eroded carbon is uncertain and might even offset LUC emissions [@wang_human-induced_2017]. Whereas for soil quality analysis SOC displacement might play an important role, in this budget approach focusing especially on the SOC debt, displaced but not emitted SOC can be treated as SOC that remains on the cropland.
Moreover, the exclusive focus on croplands ignores LUC emission on other land-use types such as pastures, rangelands and forestry. Human interventions have led to large changes in SOC stocks there as well [@sanderman_soil_2017; @friedlingstein_global_2019]. This study does not intend to be a comprehensive LUC emission analysis and acknowledges that land-use changes comes with large overall emissions.

@@

To validate our modeled SOC stocks and stock changes under management, we compare our results to default IPCC stock changes factors of 2006 [@eggleston_ipcc_2006] and their refinements in 2019 [@calvo_buendia_ipcc_2019]. Both estimates are based on measurement data for croplands (see Table \@ref(tab:SCFglo)). To allow for comparison, we aggregate our stock change factors to the four IPCC climate zones (Fig. \@ref(fig:CLIMzone)).

```{r SCFglo, echo=FALSE, results = 'asis'}
table <- readRDS("../ResultNotebooks/Output/SFC_comparison.rds")
library(xtable)
print(xtable(table$comparison_table, digit=2, label="tab:SCFglo",
              caption="$F^{\\mathrm{SCF}}$ in comparison to IPCC Tier 1 default factors: Stock change factors are in good agreement with the default values of the IPCC in general. For the tropical regions the assumptions changed notablly from the guidelines in 2006 \\citep{eggleston_ipcc_2006} to the update in 2019 \\citep{calvo_buendia_ipcc_2019}. leaving our results in very good agreement with the old default assumptions. Default assumption are given under the assumption of medium input systems, which, considering the yield gap in mainly developing regions in the tropics, might be an overestimation and decrease $F^{\\mathrm{SCF}}$ by additional 5-8 percent. Also modelled $F^{SOC}$ have increased for all climates over time."), comment=FALSE, caption.placement = "top")
```

Whereas our estimates are lower in the two tropical climate zones, temperate zone default factors are higher than Calvo Buendia et al. [-@calvo_buendia_ipcc_2019]. The tropical factors differ substantially between the IPCC guidelines in 2006 [@eggleston_ipcc_2006] and their refinement in 2019 [@calvo_buendia_ipcc_2019], with our estimates within this range.
Taking the simplified assumption, that tropical soils might suffer from insufficient C input rates (low) due to yield gaps, whereas temperate soils in developed regions might be highly managed and fertilized (high), the difference vanishes partially. Our results still show  especially in temperate dry regions --- the smallest region area-wise --- small deviations from natural SOC stocks. Considering the impact of irrigation and fertilization on carbon-poor dryland soils, even factors above 1 (see Fig. \@ref(fig:SOCmaps)(c)) may be expected.

@@

The world's SOC stock and its changes are highly uncertain, which is seen in the wide range of global SOC stock estimates for the first 30 cm of the soil profile [@batjes_harmonized_2016; @hengl_soilgrids250m_2017; @fao_global_2018; @schaphoff_lpjml4_2018-1] in Table \@ref(fig:SOCglo)).

@@

The global estimates of the total SOC stock from this study are on the lower end compared to other modeled results or data driven estimates. SoilGrids [@hengl_soilgrids250m_2017] especially stands out for their high estimation, whereas all other sources (including our study) are comparabely similar. Looking at regional results in Fig. \@ref(fig:SOCreg), our estimates turn out to be in good agreement for most of the world, with the largest deviations for boreal moist and tropical moist areas. To avoid that this bias influences our results, which originates from uncertainties in the representation of natural land, we focus on SOC changes on cropland. Pristine natural vegetated areas (like permafrost and rain forests) without human land management thus drop out in our calculation of SOC debt and stock change factors.

### 2020-12-17 SOCbudgetPaper/introduction.Rmd
update (fix citations, affilations, figure position, LPJmL4Paper to LPJmL4, sidewaytable)

Soil Organic Carbon (SOC), the amount of organic carbon stored in the Earth’s soil, is the largest terrestrial organic carbon pool. It exceeds the carbon in the atmospheric and vegetation pools multiple times ([@batjes_total_1996]). Even small changes in processes affecting SOC lead therefore to substantial shifts in the terrestrial carbon cycle and influence the amount of CO~2~ in the atmosphere ([@friedlingstein_global_2019], [@minasny_soil_2017]). The specific amount of carbon stored in soils globally is quantified with estimates ranging from 1500 to 2400 GtC for the first meter of the soil profile ([@batjes_total_1996], [@sanderman_soil_2017]).

Natural properties like climatic, biophysical, and landscape characteristics clearly play the most important roles to determine SOC variations over space and time. Recent studies have focused on the evaluation of total SOC stocks of the world as well as on the spatial disaggregation of soil properties such as SOC content ([@batjes_harmonized_2016], [@hengl_soilgrids250m_2017], [@fao_global_2018]). However, these studies often do not include human interventions, like land cover change and agricultural management, in their analysis. Compared to climatic and geological drivers, they alter terrestrial carbon pools over much shorter time scales and are currently one of most dominant drivers of SOC changes on managed land.

The anthropogenic impact can be measured by the SOC debt (also referred to as SOC component of land-use change emissions), which is the amount of organic carbon soils have lost under cultivation compared to a potential natural vegetated state. [@sanderman_soil_2017] identified the anthropogenic SOC debt for the first meter of the soil profile due to land cover change at around 116 GtC (37 GtC for the first 30 cm), compared to previous estimates of 60--130 GtC for the first meter ([@lal_world_2001]).

Global assessments of the carbon cycle via dynamic global vegetation models (DGVMs) and Earth System Models (ESMs) or bookkeeping models (BKMs) have analyzed SOC losses as part of a comprehensive evaluation of the global carbon budget and land-use change (LUC) emissions ([@friedlingstein_global_2019]). While providing estimates of the magnitude of SOC losses due to land cover change, most models lack a detailed consideration of agricultural management. Earlier DGVM and ESM based assessments only considered changes in land cover, but ignored the removal of biomass at harvest ([@strassmann_simulating_2008], [@betts_climate_2015]). BKMs are designed to estimate LUC related emissons but ignore changes in SOC due to climate change, CO~2~ fertilization, and agricultural management ([@friedlingstein_global_2019], [@houghton_carbon_2012], [@hansis_relevance_2015]).

Managed agricultural systems were introduced in greater detail to DGVMs and ESMs to improve the assessment of the terrestrial carbon balance (e.g. [@bondeau_modelling_2007], [@lindeskog_implications_2013]). [@pugh_simulated_2015] explicitly consider agricultural management in the form of tillage, irrigation and biomass extraction at harvest, but worked with stylized scenarios rather than with historic management data. They also showed the importance of accounting for the land-use history, as many carbon emissions from agricultural soils are caused by historic LUC and the slow decline of SOC under cropland before it reaches a new equilibrium.

In global-scale carbon cycle assessments, management systems are typically represented as spatially explicit patterns that are static in time (e.g. growing seasons ([portmann_mirca2000global_2010]) multiple cropping systems ([@waha_multiple_2020]), irrigation systems ([@jagermeyr_water_2015])) or as stylized scenarios (e.g. [@pugh_simulated_2015], [@lutz_simulating_2019]).

More data sets on spatially explicit agricultural management time series with global coverage become available (e.g. on tillage systems, see [@porwollik_generating_2018], [@prestele_spatially_2018]) and model approaches are increasingly developed to project the dynamics of management systems into the future (e.g. [@iizumi_modeling_2019], [@minoli_modelling_2019]), but have -- to our knowledge -- not found their way into comprehensive assessments of the terrestrial carbon cycle in DGVMs and BKMs.

Field-scale models (@del_grosso_simulated_2001], [@coleman_simulating_1997],[@smith_estimating_2010], [@taghizadeh-toosi_c-tool_2014]) are able to better account for historic agricultural management if detailed information on crop yield levels, fertilizer inputs and various other on-farm measures is available for the studied sites. However, due to the lack of comprehensive global management data as input to these models, scaling up to the global domain remains a complex challenge.

>>>>

Soil Organic Carbon (SOC), the amount of organic carbon stored in the Earth’s soil, is the largest terrestrial organic carbon pool. It exceeds the carbon in the atmospheric and vegetation pools multiple times [@batjes_total_1996]. Even small changes in processes affecting SOC lead therefore to substantial shifts in the terrestrial carbon cycle and influence the amount of CO~2~ in the atmosphere [@friedlingstein_global_2019; @minasny_soil_2017]. The specific amount of carbon stored in soils globally is quantified with estimates ranging from 1500 to 2400 GtC for the first meter of the soil profile [@batjes_total_1996; @sanderman_soil_2017].

Natural properties like climatic, biophysical, and landscape characteristics clearly play the most important roles to determine SOC variations over space and time. Recent studies have focused on the evaluation of total SOC stocks of the world as well as on the spatial disaggregation of soil properties such as SOC content [@batjes_harmonized_2016; @hengl_soilgrids250m_2017; @fao_global_2018]. However, these studies often do not include human interventions, like land cover change and agricultural management, in their analysis. Compared to climatic and geological drivers, they alter terrestrial carbon pools over much shorter time scales and are currently one of most dominant drivers of SOC changes on managed land.

The anthropogenic impact can be measured by the SOC debt (also referred to as SOC component of land-use change emissions), which is the amount of organic carbon soils have lost under cultivation compared to a potential natural vegetated state. Sanderman et al. [-@sanderman_soil_2017] identified the anthropogenic SOC debt for the first meter of the soil profile due to land cover change at around 116 GtC (37 GtC for the first 30 cm), compared to previous estimates of 60--130 GtC for the first meter [@lal_world_2001].

Global assessments of the carbon cycle via dynamic global vegetation models (DGVMs) and Earth System Models (ESMs) or bookkeeping models (BKMs) have analyzed SOC losses as part of a comprehensive evaluation of the global carbon budget and land-use change (LUC) emissions [@friedlingstein_global_2019]. While providing estimates of the magnitude of SOC losses due to land cover change, most models lack a detailed consideration of agricultural management. Earlier DGVM and ESM based assessments only considered changes in land cover, but ignored the removal of biomass at harvest [@strassmann_simulating_2008; @betts_climate_2015]. BKMs are designed to estimate LUC related emissons but ignore changes in SOC due to climate change, CO~2~ fertilization, and agricultural management [@friedlingstein_global_2019; @houghton_carbon_2012; @hansis_relevance_2015].

Managed agricultural systems were introduced in greater detail to DGVMs and ESMs to improve the assessment of the terrestrial carbon balance [e.g. @bondeau_modelling_2007; @lindeskog_implications_2013]. Pugh et al. [-@pugh_simulated_2015] explicitly consider agricultural management in the form of tillage, irrigation and biomass extraction at harvest, but worked with stylized scenarios rather than with historic management data. They also showed the importance of accounting for the land-use history, as many carbon emissions from agricultural soils are caused by historic LUC and the slow decline of SOC under cropland before it reaches a new equilibrium.

In global-scale carbon cycle assessments, management systems are typically represented as spatially explicit patterns that are static in time (e.g. growing seasons [@portmann_mirca2000global_2010] multiple cropping systems [@waha_multiple_2020], irrigation systems [@jagermeyr_water_2015]) or as stylized scenarios [e.g. @pugh_simulated_2015, @lutz_simulating_2019].

More data sets on spatially explicit agricultural management time series with global coverage become available (e.g. on tillage systems, see [@porwollik_generating_2018], [@prestele_spatially_2018]) and model approaches are increasingly developed to project the dynamics of management systems into the future (e.g. [@iizumi_modeling_2019], [@minoli_modelling_2019]), but have --- to our knowledge --- not found their way into comprehensive assessments of the terrestrial carbon cycle in DGVMs and BKMs.

Field-scale models [@del_grosso_simulated_2001, @coleman_simulating_1997, @smith_estimating_2010, @taghizadeh-toosi_c-tool_2014] are able to better account for historic agricultural management if detailed information on crop yield levels, fertilizer inputs and various other on-farm measures is available for the studied sites. However, due to the lack of comprehensive global management data as input to these models, scaling up to the global domain remains a complex challenge.
